Below-ground sink limitation alters growth and carbon balance of Eucalyptus seedlings

### 2015-10-23 manuscript/manuscript_pve.Rmd
manuscripts edits from MT

## Gas exchange and photosynthetic parameters
At the first measurement campaign, both A~sat~ and A~max~ were significantly higher in the free seedling treatment compared to seedlings in containers (both P < 0.001). Across all measurement campaigns mean A~sat~ (Figure `r asatfig`) and A~max~ (Table 2) were consistently higher in free seedlings than in containers (26 % and 29 %, respectively). The relationships between photosynthetic capacity, leaf starch, and leaf N on a mass basis was marginally significant (P = 0.058) but A~max~ on a mass basis was highly correlated to both leaf N content and leaf starch (both P < 0.001). Across all measurement campaigns and treatments A~max~ was higher when foliar N was also higher, usually associated with low levels of leaf starch (Figure `r achemfig`a). A~max~  was also lower when leaf starch was high as higher leaf N often did not coincide with high leaf starch (Figure `r achemfig`b). Overall, A~max~ was postively correlated with final harvest biomass across all seedlings (P < 0.001).

@@

# Discussion
This study utilized a simple but novel field design to manipulate belowground sink limitation and physically restrict *Eucalyptus tereticornis* seedling biomass production. We addressed questions regarding the coordination of A and growth by complementing empirical results with a C balance model. We found that reductions in leaf A across container sizes, when integrated across the 120 day experiment, were alone insufficient to account for observed reductions in total plant biomass production.

We thus encourage the utilization of mass balance approaches and provide direction for future studies when testing factors that control plant growth under environmental change.

@@

\
Using measured LMF from the harvest, instead of the optimized seedling control (M~2~), improved biomass predictions and provided insight into how sink limitation can impact leaf C allocation beyond A. The sensitivity of the model to shifts in LMF could represent changes in senescence that could not be explicitly quantified in this field study. As TNC accumulation can lead to accelerated leaf senescence [@paul2001sink], this could explain the observed decline in total leaf area of seedlings in small containers. Future empirical and modelling studies should focus on how feedbacks from sink activity affect both rates of A and the fate of C allocated to leaves. It will be the interactions between these two components that will determine the total C gain available for plant growth.

@@

## Conclusions

The debate over how rates of photosynthesis affect plant growth or to what degree these rates are instead controlled by growth has existed for decades [@sweet1966role]. Here, although biomass partitioning was conserved our model results infer that the allocation of photosynthate between non-biomass pools was altered by sink inhibition. This is important as manipulations of plants grown in containers are often used to draw conclusions about growth and physiological principles, but how these results actually reflect field-grown plants has seldom been studied. Our findings add more evidence that A and growth are not always entirely synced, an important distinction often missed in studies that manipulate source/sink activity. K&#246;rner [-@korner2013growth] suggests that it is the norm for sink activity to feedback onto source activity, causing growth to control A through the demand for C. Although this may be true, we argue that attempts to quantify or at least predict the fate of assimilated C into known pools of growth, storage and C loss are needed prior to addressing this debate. Our modelling results agree with conclusions from Valentine and M&#228;kel&#228; [-@valentine2005bridging] where the problem with predicting tree growth is a problem in forecasting the assimilation and allocation of C and other constituents. The approach used here has the flexibility to account for multiple drivers of C allocation and provides an avenue to address future questions regarding the impact of environmental change on plant growth.

@@

\
**Figure 3**. Soil volume treatment means of biomass partitioning to leaves, stems, and roots at harvest (a), bi-variate relationships between mass allocation to leaves and stems + roots (b) and leaf mass as a function of fine root biomass with &pm; standard error (c). For (b) lines represent standardized major axis fitting of the log transformed allometric relationships of leaf mass fraction by treatment. For (c) the dashed line is the 1:1 relationship and the solid line represents the significant linear model fit (R^2^=0.81).

\
**Figure 4**. Soil volume treatment means &pm; standard error, across all measurement campaigns (n=6), of light saturated rates of photosynthesis at 25&deg;C. Different letters represent significant differences between treatments.

>>>>

## Gas exchange and photosynthetic parameters
At the first measurement campaign, both A~sat~ and A~max~ were significantly higher in the free seedling treatment compared to seedlings in containers (both P < 0.001). Across all measurement campaigns mean A~sat~ (Figure `r asatfig`) and A~max~ (Table 2) were consistently higher in free seedlings than in containers (26 % and 29 %, respectively). The relationships between photosynthetic capacity, leaf starch, and leaf N on a mass basis was marginally significant (P = 0.058) but A~max~ on a mass basis was highly correlated to both leaf N content and leaf starch (both P < 0.001). We used predictions from the linear mixed effect model equation to visualize these relationship of A~max~ to either leaf N content or leaf starch at muliple bin levels (n=5) of the covariate parameter (Figure `r achemfig`). Across all measurement campaigns and treatments A~max~ was higher when foliar N was also higher, usually associated with low levels of leaf starch (Figure `r achemfig`a). A~max~  was also lower when leaf starch was high as higher leaf N often did not coincide with high leaf starch (Figure `r achemfig`b). Overall, A~max~ was postively correlated with final harvest biomass across all seedlings (P < 0.001).

@@

# Discussion
This study utilized a simple but novel field design to manipulate belowground sink limitation and physically restrict *Eucalyptus tereticornis* seedling biomass production. We addressed questions regarding the coordination of A and growth by complementing empirical results with a C mass balance model. We found that reductions in leaf A across container sizes, when integrated across the 120 day experiment, were alone insufficient to account for observed reductions in total plant biomass production.

@@

\
Using measured LMF from the harvest, instead of the optimized seedling control (M~2~), improved biomass predictions and provided insight into how sink limitation can impact leaf C allocation beyond A. The sensitivity of the model to shifts in LMF could represent leaf loss throughout the experiment that could not be explicitly quantified in this field study. As TNC accumulation can lead to accelerated leaf senescence [@paul2001sink], this could explain the observed decline in total leaf area of seedlings in small containers. Future empirical and modelling studies should focus on how feedbacks from sink activity affect both rates of A and the fate of C allocated to growth, respiration and C storage in leaves. It will be the interactions between these two components that will determine the total C gain available for plant growth.

@@

## Conclusions
The debate over how rates of photosynthesis affect plant growth or to what degree these rates are instead controlled by growth has existed for decades [@sweet1966role]. Here, although biomass partitioning was conserved our model results infer that the allocation of photosynthate between non-biomass pools was altered by sink inhibition. This is important as manipulations of plants grown in containers are often used to draw conclusions about growth and physiological principles, but how these results actually reflect field-grown plants has seldom been studied. Our findings add more evidence that A and growth are not always entirely synced, an important distinction often missed in studies that manipulate source/sink activity. K&#246;rner [-@korner2013growth] suggests that it is the norm for sink activity to feedback onto source activity, causing growth to control A through the demand for C. Although this may be true, we argue that attempts to quantify or at least predict the fate of assimilated C into known pools of growth, storage and C loss are needed prior to addressing this debate. Our modelling results agree with conclusions from Valentine and M&#228;kel&#228; [-@valentine2005bridging] where the problem with predicting tree growth is a problem in forecasting the assimilation and allocation of C and other constituents. The approach used here has the flexibility to account for multiple drivers of C allocation and provides an avenue to address future questions regarding the impact of environmental change on plant growth.

@@

\
**Figure 3**. Soil volume treatment means of biomass partitioning to leaves, stems, and roots at harvest (a), bivariate relationships between mass allocation to leaves and stems + roots (b) and leaf mass as a function of fine root biomass with &pm; standard error (c). For (b) lines represent standardized major axis fitting of the log-transformed allometric relationships of leaf mass fraction by treatment. For (c) the dashed line is the 1:1 relationship and the solid line represents the significant linear model fit (R^2^=0.81).

\
**Figure 4**. Soil volume treatment means &pm; standard error, across all measurement campaigns (n = 6), of light saturated rates of photosynthesis at 25&deg;C. Different letters represent significant differences between treatments.
